---
title: "HST"
author: "Takahiro"
date: "2019/11/5"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load dataset.

```{r}
dat <- read.csv("C:/Users/ogmcd/Dropbox/00_2019_Class/Fall/HST953/953-2019-team_f/data/hypo_dx.csv")
```


```{r}
library(skimr)
library(tidyverse)
library(psych)
library(MASS)
library(tableone)
library(mgcv)
library(gridExtra)
library(grid)
```

# Exprolatory Data Analyses
We found some missing values in our dataset mainly in the laboratory values.

```{r}
dat %>% skim
```

```{r}
dat %>% str
dat %>% describe
dat %>% head
```

```{r}
dat$unitType %>% summary
```


```{r}
dat.2 <- data.frame(dat)
```


```{r}
dat.2 <-
  mutate(dat.2, Male = if_else(gender == 1, true = "Yes", false = "No"))
dat.2$Male <- as.factor(dat.2$Male)
dat.2$Male <- relevel(dat.2$Male, ref="No")

dat.2$Primary_diagnosis <- relevel(dat.2$Primary_diagnosis, ref="Metabolic/Electrolytes")
dat.2$ethnicity <- relevel(dat.2$ethnicity, ref="Caucasian")
dat.2$unitType <- relevel(dat.2$unitType, ref="Med-Surg ICU")

dat.2$base_CHF <- as.factor(dat.2$base_CHF)
dat.2$base_CKD <- as.factor(dat.2$base_CKD)
dat.2$base_diabetes <- as.factor(dat.2$base_diabetes)
dat.2$base_HIV <- as.factor(dat.2$base_HIV)
dat.2$base_hypertension <- as.factor(dat.2$base_hypertension)
dat.2$base_liver_f <- as.factor(dat.2$base_liver_f)
dat.2$base_PAD <- as.factor(dat.2$base_PAD)
dat.2$base_sepsis <- as.factor(dat.2$base_sepsis)
dat.2$base_COPD <- as.factor(dat.2$base_COPD)
dat.2$base_stroke <- as.factor(dat.2$base_stroke)
dat.2$vent <- as.factor(dat.2$vent)

dat.2$prior_diuretics <- 0
dat.2[dat.2$had_bumetanide==1,]$prior_diuretics <- 1
dat.2[dat.2$had_furosemide==1,]$prior_diuretics <- 1
dat.2[dat.2$had_HCTZ==1,]$prior_diuretics <- 1
dat.2[dat.2$had_torsemide==1,]$prior_diuretics <- 1
dat.2$prior_diuretics <- as.factor(dat.2$prior_diuretics)

dat.2$prior_neurl_dis <- 0
dat.2[dat.2$base_coma==1,]$prior_neurl_dis <- 1
dat.2[dat.2$base_other_coma==1,]$prior_neurl_dis <- 1
dat.2[dat.2$base_other_para==1,]$prior_neurl_dis <- 1
dat.2[dat.2$base_seizure==1,]$prior_neurl_dis <- 1
dat.2$prior_neurl_dis <- as.factor(dat.2$prior_neurl_dis)

dat.2$out_neurl_dis <- 0
dat.2[dat.2$coma_disch==1,]$out_neurl_dis <- 1
dat.2[dat.2$other_coma_disch==1,]$out_neurl_dis <- 1
dat.2[dat.2$other_para_disch==1,]$out_neurl_dis <- 1
dat.2[dat.2$seizure_disch==1,]$out_neurl_dis <- 1

dat.2$prior_psy_dis <- 0
dat.2[dat.2$base_bipolar==1,]$prior_psy_dis <- 1
#dat.2[dat.2$base_scz==1,]$prior_psy_dis <- 1
dat.2$prior_psy_dis <- as.factor(dat.2$prior_psy_dis)

dat.2$prior_renal_f <- 0
dat.2[dat.2$base_renal_f==1,]$prior_renal_f <- 1
dat.2[dat.2$base_unspec_renal_f==1,]$prior_renal_f <- 1
dat.2$prior_renal_f <- as.factor(dat.2$prior_renal_f)

dat.2$ethnicity[dat.2$ethnicity=='Native American'] <- 'Other/Unknown'
dat.2$ethnicity <- dat.2$ethnicity[, drop = TRUE]

levels(dat.2$unitType) <- c(levels(dat.2$unitType), 'Cardiac')
dat.2$unitType[dat.2$unitType=='Cardiac ICU'] <- 'Cardiac'
dat.2$unitType[dat.2$unitType=='CCU-CTICU'] <- 'Cardiac'
dat.2$unitType[dat.2$unitType=='CSICU'] <- 'Cardiac'
dat.2$unitType[dat.2$unitType=='CTICU'] <- 'Cardiac'
dat.2$unitType[dat.2$unitType=='Med-Surg ICU'] <- 'SICU'
dat.2$unitType <- dat.2$unitType[, drop = TRUE]
```


```{r}
drops.2 <- c('had_bumetanide','had_furosemide', 'had_HCTZ', 'had_torsemide', 'base_coma', 'base_other_coma',
             'base_other_para', 'base_seizure', 'base_bipolar', 'base_SCZ', 'base_renal_f', 'base_unspec_renal_f',
             'coma_disch', 'other_coma_disch', 'other_para_disch', 'seizure_disch', 'prior_renal_f',
             'had_D5W', 'had_DDAVP', 'had_diuretics', 'gender', 'base_HIV', 'base_PAD', 'base_sepsis',
             'prior_psy_dis')
dat.2 <- dat.2[ , !(names(dat.2) %in% drops.2)]
```

```{r}
dat.2 %>% skim()
```

Defined exposure as s[Na]+ change >6 (mmol/L).

```{r}
dat.2 <-
  mutate(dat.2, exposure = if_else(change_over_6 == 1, true = 1, false = 0))
```

Defined hospital-free days, ICU-free days as x-free days within the first 28 days.
Patients who died in the hospital were set to be 0 for both variables.

```{r}
dat.2 <-
  mutate(dat.2, ICU_days = round(icu_los_hours/24) + 1)
dat.2$ICU_days[dat.2$ICU_days > 28] <- 28
dat.2 <-
  mutate(dat.2, ICU_free_days = 28-ICU_days)
dat.2$ICU_free_days[dat.2$hosp_mortality > 0] <- 0
dat.2$ICU_free_days %>% summary

dat.2 <-
  mutate(dat.2, hos_days = round(hos_los_hours/24) + 1)
dat.2$hos_days[dat.2$hos_days > 28] <- 28
dat.2 <-
  mutate(dat.2, hos_free_days = 28-hos_days)
dat.2$hos_free_days[dat.2$hosp_mortality > 0] <- 0
dat.2$hos_free_days %>% summary
```

Dropped unnecessary columns.

```{r}
drops.3 <- c('icu_los_hours','hos_los_hours', 'hos_days', 'ICU_days', 'unitdischargelocation', 'change_over_6')
dat.3 <- dat.2[ , !(names(dat.2) %in% drops.2)]
```


```{r}
covariates <- c('base_CHF', 'base_CKD', 'base_COPD', 'base_diabetes', ' base_hypertension', 'base_liver_f', 
                'base_stroke', 'ethnicity', 'Male','Primary_diagnosis', 'prior_diuretics', 'prior_neurl_dis',
                'unitType', 'vent', 'age', 'albumin', 'ALT', 'apachescore', 'AST', 'base_sodium',
                'bicarbonate', 'BUN', 'creatinine', 'GCS', 'glucose', 'magnesium', 'phosphate', 'potassium',
                'tbili')
outcome <- c('ICU_free_days', 'hos_free_days', 'hosp_mortality', 'out_neurl_dis')
```

```{r}
tab1 <- CreateTableOne(vars = c(covariates, outcome), strata = "exposure", data = dat.3)
print(tab1, smd=TRUE)
```

Distribution in the final dataset.

```{r}
names(dat.3)
skim(dat.3)
```

# Generalized Additive Models

First we took a look at a histogram of serum sodium level on admission.

```{r}
hist(dat.3$base_sodium)
```

We conducted the GAM to explore associations between sodium increase and primary and secondary outcomes in the entire dataset.

## GAMs
Generalized additive models using initial serum sodium and age as continuous variables and sex as categorical variable.
We obtained predicted values from GAMs.

```{r}
get.predict.df <- function(df){
  
  # Create grid for prediction
  sod.dif.seq <- seq(0, 25, length=251)
  predict.df <- data.frame(sodium_change=sod.dif.seq, age=48, Male="No")
  
  # Generalized additive model for in-hospital mortality
  gam.hosp <- gam(hosp_mortality ~ s(sodium_change) + s(age) + Male,
                  data = df, family=binomial(link='logit'), method='REML')
  
  # Obtain predicted in-hospital mortality and 95% CI for a 48-year-old female
  preds.hosp <- predict(gam.hosp, type="link", newdata=predict.df, se.fit=TRUE)
  fit.hosp <- preds.hosp$fit
  fit.hosp.low95 <- fit.hosp - 1.96 * preds.hosp$se.fit
  fit.hosp.up95 <- fit.hosp + 1.96 * preds.hosp$se.fit
  
  # Convert Log odds to probability
  predict.df$hosp <- 100*(exp(fit.hosp)/(1 + exp(fit.hosp)))
  predict.df$hosp.low95 <- 100*(exp(fit.hosp.low95)/(1 + exp(fit.hosp.low95)))
  predict.df$hosp.up95 <- 100*(exp(fit.hosp.up95)/(1 + exp(fit.hosp.up95)))
  
  # Generalized additive model for neurological complication
  gam.neur <- gam(out_neurl_dis ~ s(sodium_change) + s(age) + Male,
                  data = df, family=binomial(link='logit'), method='REML')
  
  # Obtain predicted neurological complication rate and 95% CI for a 48-year-old female
  preds.neur <- predict(gam.neur, type="link", newdata=predict.df, se.fit=TRUE)
  fit.neur <- preds.neur$fit
  fit.neur.low95 <- fit.neur - 1.96 * preds.neur$se.fit
  fit.neur.up95 <- fit.neur + 1.96 * preds.neur$se.fit
  
  # Convert Log odds to probability
  predict.df$neur <- 100*(exp(fit.neur)/(1 + exp(fit.neur)))
  predict.df$neur.low95 <- 100*(exp(fit.neur.low95)/(1 + exp(fit.neur.low95)))
  predict.df$neur.up95 <- 100*(exp(fit.neur.up95)/(1 + exp(fit.neur.up95)))
  
  # Generalized additive model for hospital-free days
  gam.hos.fr <- gam(hos_free_days ~ s(sodium_change) + s(age) + Male,
                    data = df, method='REML')

  # Obtain predicted hospital-free days and 95% CI for a 48-year-old female
  preds.hos.fr <- predict(gam.hos.fr, type="link", newdata=predict.df, se.fit=TRUE)
  predict.df$hos.fr <- preds.hos.fr$fit
  predict.df$hos.fr.low95 <- preds.hos.fr$fit - 1.96 * preds.hos.fr$se.fit
  predict.df$hos.fr.up95 <- preds.hos.fr$fit + 1.96 * preds.hos.fr$se.fit
  
  # Generalized additive model for ICU-free days
  gam.ICU.fr <- gam(ICU_free_days ~ s(sodium_change) + s(age) + Male,
                    data = df, method='REML')

  # Obtain predicted ICU-free days and 95% CI for a 48-year-old female
  preds.ICU.fr <- predict(gam.ICU.fr, type="link", newdata=predict.df, se.fit=TRUE)
  predict.df$ICU.fr <- preds.ICU.fr$fit
  predict.df$ICU.fr.low95 <- preds.ICU.fr$fit - 1.96 * preds.ICU.fr$se.fit
  predict.df$ICU.fr.up95 <- preds.ICU.fr$fit + 1.96 * preds.ICU.fr$se.fit
  
  return(predict.df)
}
```


```{r}
predict.all <- get.predict.df(dat.3)
```

## Plots

```{r}
get.four.plots <- function(df){
  hosp.spline <- ggplot(df, aes(x = sodium_change, y = hosp)) +
  geom_ribbon(aes(ymin = hosp.low95,
                  ymax = hosp.up95,
                  fill = hosp.up95 > hosp.low95),
              alpha = 0.5) +
  geom_line() +
  scale_fill_manual(values="grey") +
  guides(fill = FALSE) +
  theme_bw() +
  xlim(0, 25) +
  ylim(0, 60) +
  xlab('Serum [Na+] correction rate (mmol/L/day)') +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.text.x = element_text(size = 12)) +
  ylab('In-hospital mortality (%)') +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text.y = element_text(size = 12))
  
  neur.spline <- ggplot(df, aes(x = sodium_change, y = neur)) +
  geom_ribbon(aes(ymin = neur.low95,
                  ymax = neur.up95,
                  fill = neur.up95 > neur.low95),
              alpha = 0.5) +
  geom_line() +
  scale_fill_manual(values="grey") +
  guides(fill = FALSE) +
  theme_bw() +
  xlim(0, 25) +
  ylim(0, 60) +
  xlab('Serum [Na+] correction rate (mmol/L/day)') +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.text.x = element_text(size = 12)) +
  ylab('Neurological complication rate (%)') +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text.y = element_text(size = 12))
  
  hos.fr.spline <- ggplot(df, aes(x = sodium_change, y = hos.fr)) +
  geom_ribbon(aes(ymin = hos.fr.low95,
                  ymax = hos.fr.up95,
                  fill = hos.fr.up95 > hos.fr.low95),
              alpha = 0.5) +
  geom_line() +
  scale_fill_manual(values="grey") +
  guides(fill = FALSE) +
  theme_bw() +
  xlim(0, 25) +
  xlab('Serum [Na+] correction rate (mmol/L/day)') +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.text.x = element_text(size = 12)) +
  scale_y_continuous(breaks = seq(0, 28, 7), limits = c(0, 28)) +
  ylab('Hospital-free days (day)') +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text.y = element_text(size = 12))
  
  ICU.fr.spline <- ggplot(df, aes(x = sodium_change, y = ICU.fr)) +
  geom_ribbon(aes(ymin = ICU.fr.low95,
                  ymax = ICU.fr.up95,
                  fill = ICU.fr.up95 > ICU.fr.low95),
              alpha = 0.5) +
  geom_line() +
  scale_fill_manual(values="grey") +
  guides(fill = FALSE) +
  theme_bw() +
  xlim(0, 25) +
  xlab('Serum [Na+] correction rate (mmol/L/day)') +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.text.x = element_text(size = 12)) +
  scale_y_continuous(breaks = seq(0, 28, 7), limits = c(0, 28)) +
  ylab('ICU-free days (day)') +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text.y = element_text(size = 12))
  
  return(list(hosp.spline, neur.spline, hos.fr.spline, ICU.fr.spline))
}
```


```{r}
all.plots <- get.four.plots(predict.all)

all.spline <- grid.arrange(
  all.plots[[1]],
  all.plots[[2]],
  all.plots[[3]],
  all.plots[[4]],
  nrow = 2,
  top=textGrob("Entire Population (n=1024)", gp=gpar(fontsize=18))
)

all.spline

ggsave("All_spline.png", all.spline, height = 8 , width = 12, dpi=600)
```


## Multiple Imputation
We performed multiple imputation using 'mice' package. For continuous variables, we used Bayesian linear regression. For binary variable we used logistic regression.
Imputation was conducted 100 times. Max iteration was set as 50.

```{r}
library(mice)
library(miceadds)
```

```{r}
imp.meth = c('', '', '', '', '', 'logreg', 'norm', 'norm', '', '', '', 'norm', 'norm', 'norm', 'norm', 'norm', 
             'norm', 'norm', '', 'norm', 'norm', 'norm', '', '', '', '', '', '', '', '', '', '', '', 
             '', '', '', '', '', '', '', '', '')
dat.imp <- mice(dat.3, m=100, maxit=50, meth=imp.meth, seed=953)
```

```{r}
imp.1 <- complete(dat.imp,1)
names(imp.1)
```

Note: This chunk saves all imputed datasets to csv files.

```{r}
for (i in 1:100)
{
   write.csv(complete(dat.imp, i), file=paste("imp", i, ".csv", sep=""))
}
```

# Analyses in each dataset
## Marginal Structural Model of Inverse Probability of Treatment Weighting
## Parametric g-formula
## Outcome Regression Conditional on Propensity Score

We conducted IPTW to obtain marginal average treatment effect.
For sensitivity analyses, we also performed g-formula and logistic/linear regression conditioning on propensity score.
Point estimates and variances of average treatment effects (marginal or conditional on propensity score) were obtained using 1000 times bootstrap for each imputed dataset.
We pooled the means and variances of average treatment effects across the 100 imputed datasets and estimated average treatment effects and their 95% confidence intervals.
Thus, we used linear regression regardless of the types of outcomes.


```{r}
length(imp.1$base_sodium[imp.1$base_sodium>115])
length(imp.1$base_sodium[imp.1$base_sodium<=115])
```
```{r}
psmod<-glm(exposure ~ Primary_diagnosis + age + ethnicity + unitType + vent
                     + GCS + apachescore + glucose + BUN + AST + tbili + magnesium + potassium
                     + phosphate + albumin + bicarbonate + base_hypertension + base_CKD + base_CHF + base_liver_f
                     + base_diabetes + base_COPD + base_stroke + base_sodium + Male + prior_diuretics
                     + prior_neurl_dis + base_sodium:base_CHF + base_sodium:base_CKD
                     + base_sodium:base_liver_f + base_sodium:base_diabetes + base_sodium:base_stroke
                     + base_sodium:prior_neurl_dis,
           data=imp.1, family="binomial")
imp.1$pscore <- fitted(psmod)

  # Create Weights
  
imp.1$IPTW<-ifelse(imp.1$exposure==1, 1/imp.1$pscore, 1/(1-imp.1$pscore))

EMM.hosp <- lm(hosp_mortality ~ exposure*severe, data=imp.1, weights=IPTW)

EMM.hosp %>% summary

EMM.hosp$coefficients[4]

```



```{r}
library(modelr)
```

First, we created a function to bootstrap n_iter times and conduct three main analyses and one subgroup analysis in each resampled dataset.

```{r}
get.boot.results <- function(df, n_iter) {
  
df <-
  mutate(df, severe = if_else(base_sodium > 115, true = 0, false = 1))  

  
  RD.hosp.i <- as.vector(NULL); RD.neur.i <- as.vector(NULL); RD.hosfree.i <- as.vector(NULL); RD.ICUfree.i <- as.vector(NULL)
  
  RD.hosp.p <- as.vector(NULL); RD.neur.p <- as.vector(NULL); RD.hosfree.p <- as.vector(NULL); RD.ICUfree.p <- as.vector(NULL)
  
  RD.hosp.g <- as.vector(NULL); RD.neur.g <- as.vector(NULL); RD.hosfree.g <- as.vector(NULL); RD.ICUfree.g <- as.vector(NULL)
  
  EMM.hosp.i <- as.vector(NULL); EMM.neur.i <- as.vector(NULL); EMM.hosfree.i <- as.vector(NULL); EMM.ICUfree.i <- as.vector(NULL)


for (i in 1:n_iter){
  
  # Bootstrap resample
  
  resample.data <- resample_bootstrap(df)
  resample.df <- as.data.frame(resample.data)
  
  
  # Calculate Propensity Score
  
  psmod.re<-glm(exposure ~ Primary_diagnosis + age + ethnicity + unitType + vent
                     + GCS + apachescore + glucose + BUN + AST + tbili + magnesium + potassium
                     + phosphate + albumin + bicarbonate + base_hypertension + base_CKD + base_CHF + base_liver_f
                     + base_diabetes + base_COPD + base_stroke + base_sodium + Male + prior_diuretics
                     + prior_neurl_dis + base_sodium:base_CHF + base_sodium:base_CKD
                     + base_sodium:base_liver_f + base_sodium:base_diabetes + base_sodium:base_stroke
                     + base_sodium:prior_neurl_dis,
           data=resample.df, family="binomial")
  resample.df$pscore <- fitted(psmod.re)

  # Create Weights
  
  resample.df$IPTW<-ifelse(resample.df$exposure==1, 1/resample.df$pscore, 1/(1-resample.df$pscore))

  # Fit MSM
  
  msm.hosp.re <- lm(hosp_mortality ~ exposure, data=resample.df, weights=IPTW)
  RD.hosp.i[[i]] <- msm.hosp.re$coefficients[2]

  msm.neur.re <- lm(out_neurl_dis ~ exposure, data=resample.df, weights=IPTW)
  RD.neur.i[[i]] <- msm.neur.re$coefficients[2]

  msm.hosfree.re <- lm(hos_free_days ~ exposure, data=resample.df, weights=IPTW)
  RD.hosfree.i[[i]] <- msm.hosfree.re$coefficients[2]

  msm.ICUfree.re <- lm(ICU_free_days ~ exposure, data=resample.df, weights=IPTW)
  RD.ICUfree.i[[i]] <- msm.ICUfree.re$coefficients[2]
  
  # PS model
  
  resample.df$dec <- cut(resample.df$pscore, 
                        breaks=c(quantile(resample.df$pscore, probs=seq(0,1,0.1))),
                        include.lowest=TRUE)

  # Fit outcome regression model
  
  out.hosp.re <- lm(hosp_mortality ~ exposure + as.factor(dec), data=resample.df)
  RD.hosp.p[[i]] <- out.hosp.re$coefficients[2]

  out.neur.re <- lm(out_neurl_dis ~ exposure + as.factor(dec), data=resample.df)
  RD.neur.p[[i]] <- out.neur.re$coefficients[2]

  out.hosfree.re <- lm(hos_free_days ~ exposure + as.factor(dec), data=resample.df)
  RD.hosfree.p[[i]] <- out.hosfree.re$coefficients[2]

  out.ICUfree.re <- lm(ICU_free_days ~ exposure + as.factor(dec), data=resample.df)
  RD.ICUfree.p[[i]] <- out.ICUfree.re$coefficients[2]
  
  
  # g-formula
  
  resample.df$interv <- -1
  d0 <- resample.df # 2nd copy: treatment set to 0, outcome to missing
  d0$interv <- 0
  d0$exposure <- 0
  d0$hosp_mortality <- NA
  d0$out_neurl_dis <- NA
  d0$hos_free_days <- NA
  d0$ICU_free_days <- NA
  d1 <- resample.df # 3rd copy: treatment set to 1, outcome to missing
  d1$interv <- 1
  d1$exposure <- 1
  d1$hosp_mortality <- NA
  d1$out_neurl_dis <- NA
  d1$hos_free_days <- NA
  d1$ICU_free_days <- NA
  d.onesample <- rbind(resample.df, d0, d1)
  
  g.hosp.re <- glm(hosp_mortality ~ exposure + age + vent + GCS + apachescore + BUN + potassium + albumin
                   + base_CHF + base_sodium + prior_neurl_dis, data=d.onesample, family=binomial)
  
  g.neur.re <- glm(out_neurl_dis ~ exposure + age + vent + GCS + apachescore + BUN + potassium + albumin
                   + base_CHF + base_sodium + prior_neurl_dis, data=d.onesample, family=binomial)
  

  g.hosfree.re <- lm(hos_free_days ~ exposure + Primary_diagnosis + age + ethnicity + unitType + vent
                     + GCS + apachescore + glucose + BUN + AST + tbili + magnesium + potassium
                     + phosphate + albumin + bicarbonate + base_hypertension + base_CKD + base_CHF + base_liver_f
                     + base_diabetes + base_COPD + base_stroke + base_sodium + Male + prior_diuretics
                     + prior_neurl_dis + base_sodium:base_CHF + base_sodium:base_CKD
                     + base_sodium:base_liver_f + base_sodium:base_diabetes + base_sodium:base_stroke
                     + base_sodium:prior_neurl_dis,
                     data=d.onesample)
  
  g.ICUfree.re <- lm(ICU_free_days ~ exposure + Primary_diagnosis + age + ethnicity + unitType + vent
                     + GCS + apachescore + glucose + BUN + AST + tbili + magnesium + potassium
                     + phosphate + albumin + bicarbonate + base_hypertension + base_CKD + base_CHF + base_liver_f
                     + base_diabetes + base_COPD + base_stroke + base_sodium + Male + prior_diuretics
                     + prior_neurl_dis + base_sodium:base_CHF + base_sodium:base_CKD
                     + base_sodium:base_liver_f + base_sodium:base_diabetes + base_sodium:base_stroke
                     + base_sodium:prior_neurl_dis,
                     data=d.onesample)
  
  d.onesample$prob.hosp <- predict(g.hosp.re, d.onesample, type="response")
  d.onesample$prob.neur <- predict(g.neur.re, d.onesample, type="response")
  d.onesample$hosfree <- predict(g.hosfree.re, d.onesample)
  d.onesample$ICUfree <- predict(g.ICUfree.re, d.onesample)

  RD.hosp.g[[i]] <- mean(d.onesample$prob.hosp[d.onesample$interv==1])-
    mean(d.onesample$prob.hosp[d.onesample$interv==0])
  RD.neur.g[[i]] <- mean(d.onesample$prob.neur[d.onesample$interv==1])-
    mean(d.onesample$prob.neur[d.onesample$interv==0])
  RD.hosfree.g[[i]] <- mean(d.onesample$hosfree[d.onesample$interv==1])-
    mean(d.onesample$hosfree[d.onesample$interv==0])
  RD.ICUfree.g[[i]] <- mean(d.onesample$ICUfree[d.onesample$interv==1])-
    mean(d.onesample$ICUfree[d.onesample$interv==0])
  
  
  #EMM based on IP weighting
  EMM.hosp.re <- lm(hosp_mortality ~ exposure*severe, data=resample.df, weights=IPTW)
  EMM.hosp.i[[i]] <- EMM.hosp.re$coefficients[4]

  EMM.neur.re <- lm(out_neurl_dis ~ exposure*severe, data=resample.df, weights=IPTW)
  EMM.neur.i[[i]] <- EMM.neur.re$coefficients[4]

  EMM.hosfree.re <- lm(hos_free_days ~ exposure*severe, data=resample.df, weights=IPTW)
  EMM.hosfree.i[[i]] <- EMM.hosfree.re$coefficients[4]

  EMM.ICUfree.re <- lm(ICU_free_days ~ exposure*severe, data=resample.df, weights=IPTW)
  EMM.ICUfree.i[[i]] <- EMM.ICUfree.re$coefficients[4]

}
  
# Store means and variances of outcomes to dataframe object
  
results <- data.frame(RD.hosp.i.mean=mean(RD.hosp.i), RD.hosp.i.var=var(RD.hosp.i),
                      RD.neur.i.mean=mean(RD.neur.i), RD.neur.i.var=var(RD.neur.i), 
                      RD.hosfree.i.mean=mean(RD.hosfree.i), RD.hosfree.i.var=var(RD.hosfree.i), 
                      RD.ICUfree.i.mean=mean(RD.ICUfree.i), RD.ICUfree.i.var=var(RD.ICUfree.i),
                      RD.hosp.p.mean=mean(RD.hosp.p), RD.hosp.p.var=var(RD.hosp.p),
                      RD.neur.p.mean=mean(RD.neur.p), RD.neur.p.var=var(RD.neur.p), 
                      RD.hosfree.p.mean=mean(RD.hosfree.p), RD.hosfree.p.var=var(RD.hosfree.p), 
                      RD.ICUfree.p.mean=mean(RD.ICUfree.p), RD.ICUfree.p.var=var(RD.ICUfree.p),
                      RD.hosp.g.mean=mean(RD.hosp.g), RD.hosp.g.var=var(RD.hosp.g),
                      RD.neur.g.mean=mean(RD.neur.g), RD.neur.g.var=var(RD.neur.g), 
                      RD.hosfree.g.mean=mean(RD.hosfree.g), RD.hosfree.g.var=var(RD.hosfree.g), 
                      RD.ICUfree.g.mean=mean(RD.ICUfree.g), RD.ICUfree.g.var=var(RD.ICUfree.g),
                      EMM.hosp.i.mean=mean(EMM.hosp.i), EMM.hosp.i.var=var(EMM.hosp.i),
                      EMM.neur.i.mean=mean(EMM.neur.i), EMM.neur.i.var=var(EMM.neur.i),
                      EMM.hosfree.i.mean=mean(EMM.hosfree.i), EMM.hosfree.i.var=var(EMM.hosfree.i), 
                      EMM.ICUfree.i.mean=mean(EMM.ICUfree.i), EMM.ICUfree.i.var=var(EMM.ICUfree.i))
return(results)
}
```


Actual 1000 times bootstrap for 100 imputed datasets.
Note: This chunk may take few hours.

```{r}
set.seed(953)
time1 <- Sys.time()

n_imputation <- 100

RD.hosp.i.mean.imp <- rep(NA, n_imputation); RD.neur.i.mean.imp <- rep(NA, n_imputation); RD.hosfree.i.mean.imp <- rep(NA, n_imputation); RD.ICUfree.i.mean.imp <- rep(NA, n_imputation)

RD.hosp.i.var.imp <- rep(NA, n_imputation); RD.neur.i.var.imp <- rep(NA, n_imputation); RD.hosfree.i.var.imp <- rep(NA, n_imputation); RD.ICUfree.i.var.imp <- rep(NA, n_imputation)

RD.hosp.p.mean.imp <- rep(NA, n_imputation); RD.neur.p.mean.imp <- rep(NA, n_imputation); RD.hosfree.p.mean.imp <- rep(NA, n_imputation); RD.ICUfree.p.mean.imp <- rep(NA, n_imputation)

RD.hosp.p.var.imp <- rep(NA, n_imputation); RD.neur.p.var.imp <- rep(NA, n_imputation); RD.hosfree.p.var.imp <- rep(NA, n_imputation); RD.ICUfree.p.var.imp <- rep(NA, n_imputation)
  
RD.hosp.g.mean.imp <- rep(NA, n_imputation); RD.neur.g.mean.imp <- rep(NA, n_imputation); RD.hosfree.g.mean.imp <- rep(NA, n_imputation); RD.ICUfree.g.mean.imp <- rep(NA, n_imputation)

RD.hosp.g.var.imp <- rep(NA, n_imputation); RD.neur.g.var.imp <- rep(NA, n_imputation); RD.hosfree.g.var.imp <- rep(NA, n_imputation); RD.ICUfree.g.var.imp <- rep(NA, n_imputation)

EMM.hosp.i.mean.imp <- rep(NA, n_imputation); EMM.neur.i.mean.imp <- rep(NA, n_imputation); EMM.hosfree.i.mean.imp <- rep(NA, n_imputation); EMM.ICUfree.i.mean.imp <- rep(NA, n_imputation)

EMM.hosp.i.var.imp <- rep(NA, n_imputation); EMM.neur.i.var.imp <- rep(NA, n_imputation); EMM.hosfree.i.var.imp <- rep(NA, n_imputation); EMM.ICUfree.i.var.imp <- rep(NA, n_imputation)

for (i in 1:n_imputation){
  imp <- complete(dat.imp,i)
  results.imp <- get.boot.results(imp, 1000)
  
  RD.hosp.i.mean.imp[i] <- results.imp[[1]]
  RD.hosp.i.var.imp[i] <- results.imp[[2]]
  RD.neur.i.mean.imp[i] <- results.imp[[3]]
  RD.neur.i.var.imp[i] <- results.imp[[4]]
  RD.hosfree.i.mean.imp[i] <- results.imp[[5]]
  RD.hosfree.i.var.imp[i] <- results.imp[[6]]
  RD.ICUfree.i.mean.imp[i] <- results.imp[[7]]
  RD.ICUfree.i.var.imp[i] <- results.imp[[8]]
  
  RD.hosp.p.mean.imp[i] <- results.imp[[9]]
  RD.hosp.p.var.imp[i] <- results.imp[[10]]
  RD.neur.p.mean.imp[i] <- results.imp[[11]]
  RD.neur.p.var.imp[i] <- results.imp[[12]]
  RD.hosfree.p.mean.imp[i] <- results.imp[[13]]
  RD.hosfree.p.var.imp[i] <- results.imp[[14]]
  RD.ICUfree.p.mean.imp[i] <- results.imp[[15]]
  RD.ICUfree.p.var.imp[i] <- results.imp[[16]]
  
  RD.hosp.g.mean.imp[i] <- results.imp[[17]]
  RD.hosp.g.var.imp[i] <- results.imp[[18]]
  RD.neur.g.mean.imp[i] <- results.imp[[19]]
  RD.neur.g.var.imp[i] <- results.imp[[20]]
  RD.hosfree.g.mean.imp[i] <- results.imp[[21]]
  RD.hosfree.g.var.imp[i] <- results.imp[[22]]
  RD.ICUfree.g.mean.imp[i] <- results.imp[[23]]
  RD.ICUfree.g.var.imp[i] <- results.imp[[24]]
  
  EMM.hosp.i.mean.imp[i] <- results.imp[[25]]
  EMM.hosp.i.var.imp[i] <- results.imp[[26]]
  EMM.neur.i.mean.imp[i] <- results.imp[[27]]
  EMM.neur.i.var.imp[i] <- results.imp[[28]]
  EMM.hosfree.i.mean.imp[i] <- results.imp[[29]]
  EMM.hosfree.i.var.imp[i] <- results.imp[[30]]
  EMM.ICUfree.i.mean.imp[i] <- results.imp[[31]]
  EMM.ICUfree.i.var.imp[i] <- results.imp[[32]]
}

time2 <- Sys.time()
difftime(time2, time1, units='mins')
```

Function to obtain pooled point estimates and confidence intervals using Rubin's rule.

```{r}
get.mean.confint <- function(mean.imp, var.imp, n_imputation){
  ATE <- mean(mean.imp)
  WM <- mean(var.imp)
  sigma <- as.vector(NULL)
  for (i in 1:n_imputation){
    sigma[[i]] <- (mean.imp[[i]] - ATE)^2
  }
  BM <- sum(sigma)/(n_imputation - 1)
  TM <- WM + (1+(1/n_imputation))*BM
  se <- sqrt(TM)
  ll <- ATE - qnorm(0.975)*se
  ul <- ATE + qnorm(0.975)*se
  result.df <- data.frame(ATE, ll, ul, se)
  return(result.df)
}
```

Obtain results.

```{r}
result.hosp.i <- get.mean.confint(RD.hosp.i.mean.imp, RD.hosp.i.var.imp, n_imputation)
result.hosp.p <- get.mean.confint(RD.hosp.p.mean.imp, RD.hosp.p.var.imp, n_imputation)
result.hosp.g <- get.mean.confint(RD.hosp.g.mean.imp, RD.hosp.g.var.imp, n_imputation)
result.hosp.emm <- get.mean.confint(EMM.hosp.i.mean.imp, EMM.hosp.i.var.imp, n_imputation)

result.neur.i <- get.mean.confint(RD.neur.i.mean.imp, RD.neur.i.var.imp, n_imputation)
result.neur.p <- get.mean.confint(RD.neur.p.mean.imp, RD.neur.p.var.imp, n_imputation)
result.neur.g <- get.mean.confint(RD.neur.g.mean.imp, RD.neur.g.var.imp, n_imputation)
result.neur.emm <- get.mean.confint(EMM.neur.i.mean.imp, EMM.neur.i.var.imp, n_imputation)

result.hosfree.i <- get.mean.confint(RD.hosfree.i.mean.imp, RD.hosfree.i.var.imp, n_imputation)
result.hosfree.p <- get.mean.confint(RD.hosfree.p.mean.imp, RD.hosfree.p.var.imp, n_imputation)
result.hosfree.g <- get.mean.confint(RD.hosfree.g.mean.imp, RD.hosfree.g.var.imp, n_imputation)
result.hosfree.emm <- get.mean.confint(EMM.hosfree.i.mean.imp, EMM.hosfree.i.var.imp, n_imputation)

result.ICUfree.i <- get.mean.confint(RD.ICUfree.i.mean.imp, RD.ICUfree.i.var.imp, n_imputation)
result.ICUfree.p <- get.mean.confint(RD.ICUfree.p.mean.imp, RD.ICUfree.p.var.imp, n_imputation)
result.ICUfree.g <- get.mean.confint(RD.ICUfree.g.mean.imp, RD.ICUfree.g.var.imp, n_imputation)
result.ICUfree.emm <- get.mean.confint(EMM.ICUfree.i.mean.imp, EMM.ICUfree.i.var.imp, n_imputation)
```

Note: Obtained values are point estimate, lower 95% confidence interval, upper 95% confidence interval, and standard error.

```{r}
result.hosp.i; result.hosp.p; result.hosp.g
```

```{r}
result.neur.i; result.neur.p; result.neur.g
```

```{r}
result.hosfree.i; result.hosfree.p; result.hosfree.g
```

```{r}
result.ICUfree.i; result.ICUfree.p; result.ICUfree.g
```


```{r}
result.hosp.emm
chi.hosp.emm <- (result.hosp.emm[[1]]/result.hosp.emm[[4]])^2

result.neur.emm
chi.neur.emm <- (result.neur.emm[[1]]/result.neur.emm[[4]])^2

result.hosfree.emm
chi.hosfree.emm <- (result.hosfree.emm[[1]]/result.hosfree.emm[[4]])^2

result.ICUfree.emm
chi.ICUfree.emm <- (result.ICUfree.emm[[1]]/result.ICUfree.emm[[4]])^2
```


```{r}
p_val.hosp <- pchisq(chi.hosp.emm, df = 1, lower.tail = FALSE); p_val.hosp

p_val.neur <- pchisq(chi.neur.emm, df = 1, lower.tail = FALSE); p_val.neur

p_val.hosfree <- pchisq(chi.hosfree.emm, df = 1, lower.tail = FALSE); p_val.hosfree

p_val.ICUfree <- pchisq(chi.ICUfree.emm, df = 1, lower.tail = FALSE); p_val.ICUfree
```


# Create Forest Plots

```{r}
method.hosp <- c('IP Weighting', 'Ps adjustment', 'g-formula')
beta.hosp <- c(-3.01, -2.74, -3.28)
Lower.limit.hosp <- c(-7.07, -6.50, -6.98)
Upper.limit.hosp <- c(1.05, 1.02, 0.42)
df.hosp <- data.frame(Method=method.hosp, beta=beta.hosp,
                        Lower_limit=Lower.limit.hosp, Upper_limit=Upper.limit.hosp)

df.hosp$Method <- relevel(df.hosp$Method, ref='Ps adjustment')
```


```{r}
hosp.plot = ggplot(data=df.hosp,
    aes(x = Method,y = beta, ymin = Lower_limit, ymax = Upper_limit ))+
    geom_pointrange(aes(col=Method))+
    geom_hline(aes(fill=Method),yintercept =0, linetype=2)+
    xlab("")+
    ylim(-8, 8)+
    geom_errorbar(aes(ymin=Lower_limit, ymax=Upper_limit,col=Method),width=0.2,cex=1)+ 
    theme(plot.title=element_text(size=16, color='black'),
        axis.text.y=element_text(size=16, color='black'),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(size=16, color='black'),
        axis.title=element_blank(),
        legend.position = 'none',
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180, color='black'))+
    coord_flip()+
    labs(title="In-hospital Mortality (%)")+
  theme(plot.title = element_text(hjust=0.5))
hosp.plot

ggsave("Forest_plot_hosp.png", hosp.plot, dpi=600)
```


```{r}
method.neur <- c('IP Weighting', 'Ps adjustment', 'g-formula')
beta.neur <- c(1.41, 0.70, 0.79)
Lower.limit.neur <- c(-2.12, -2.47, -2.21)
Upper.limit.neur <- c(4.95, 3.86, 3.79)
df.neur <- data.frame(Method=method.neur, beta=beta.neur,
                        Lower_limit=Lower.limit.neur, Upper_limit=Upper.limit.neur)

df.neur$Method <- relevel(df.neur$Method, ref='Ps adjustment')
```

```{r}
neur.plot = ggplot(data=df.neur,
    aes(x = Method,y = beta, ymin = Lower_limit, ymax = Upper_limit ))+
    geom_pointrange(aes(col=Method))+
    geom_hline(aes(fill=Method),yintercept =0, linetype=2)+
    xlab("")+
    ylim(-8, 8)+
    geom_errorbar(aes(ymin=Lower_limit, ymax=Upper_limit,col=Method),width=0.2,cex=1)+ 
    theme(plot.title=element_text(size=16, color='black'),
        axis.text.y=element_text(size=16, color='black'),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(size=16, color='black'),
        axis.title=element_blank(),
        legend.position = 'none',
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180, color='black'))+
    coord_flip()+
    labs(title="Neuroligcal Complication Rate (%)")+
  theme(plot.title = element_text(hjust=0.5))
neur.plot

ggsave("Forest_plot_neur.png", neur.plot, dpi=600)
```


```{r}
method.hos.fr <- c('IP Weighting', 'Ps adjustment', 'g-formula')
beta.hos.fr <- c(1.01, 0.97, 0.95)
Lower.limit.hos.fr <- c(0.04, 0.06, 0.05)
Upper.limit.hos.fr <- c(1.98, 1.89, 1.85)
df.hos.fr <- data.frame(Method=method.hos.fr, beta=beta.hos.fr,
                        Lower_limit=Lower.limit.hos.fr, Upper_limit=Upper.limit.hos.fr)

df.hos.fr$Method <- relevel(df.hos.fr$Method, ref='Ps adjustment')
```

```{r}
hos.fr.plot = ggplot(data=df.hos.fr,
    aes(x = Method,y = beta, ymin = Lower_limit, ymax = Upper_limit ))+
    geom_pointrange(aes(col=Method))+
    geom_hline(aes(fill=Method),yintercept =0, linetype=2)+
    xlab("")+
    ylim(-3, 3)+
    geom_errorbar(aes(ymin=Lower_limit, ymax=Upper_limit,col=Method),width=0.2,cex=1)+ 
    theme(plot.title=element_text(size=16, color='black'),
        axis.text.y=element_text(size=16, color='black'),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(size=16, color='black'),
        axis.title=element_blank(),
        legend.position = 'none',
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180, color='black'))+
    coord_flip()+
    labs(title="Hospital-Free Days (day)")+
  theme(plot.title = element_text(hjust=0.5))
hos.fr.plot

ggsave("Forest_plot_hosp_fr.png", hos.fr.plot, dpi=600)
```


```{r}
method.ICU.fr <- c('IP Weighting', 'Ps adjustment', 'g-formula')
beta.ICU.fr <- c(0.67, 0.62, 0.61)
Lower.limit.ICU.fr <- c(-0.35, -0.32, -0.32)
Upper.limit.ICU.fr <- c(1.69, 1.56, 1.53)
df.ICU.fr <- data.frame(Method=method.ICU.fr, beta=beta.ICU.fr,
                        Lower_limit=Lower.limit.ICU.fr, Upper_limit=Upper.limit.ICU.fr)

df.ICU.fr$Method <- relevel(df.ICU.fr$Method, ref='Ps adjustment')
```

```{r}
ICU.fr.plot = ggplot(data=df.ICU.fr,
    aes(x = Method,y = beta, ymin = Lower_limit, ymax = Upper_limit ))+
    geom_pointrange(aes(col=Method))+
    geom_hline(aes(fill=Method),yintercept =0, linetype=2)+
    xlab("")+
    ylim(-3, 3)+
    geom_errorbar(aes(ymin=Lower_limit, ymax=Upper_limit,col=Method),width=0.2,cex=1)+ 
    theme(plot.title=element_text(size=16, color='black'),
        axis.text.y=element_text(size=16, color='black'),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(size=16, color='black'),
        axis.title=element_blank(),
        legend.position = 'none',
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180, color='black'))+
    coord_flip()+
    labs(title="ICU-Free Days (day)")+
  theme(plot.title = element_text(hjust=0.5))
ICU.fr.plot

ggsave("Forest_plot_icu_fr.png", ICU.fr.plot, dpi=600)
```

